### 2014-04-23 paper/paper.tex
minor bugfixes

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09

\title{Formatting Instructions for NIPS 2014}

\author{
Stephanie deWet \\
Department of Computer Sciences\\
University of Wisconsin - Madison \\
Madison, WI 53705 \\
\texttt{sdewet@cs.wisc.edu} \\
\And
Saswati De \\
Department of Computer Sciences\\
University of Wisconsin - Madison \\
Madison, WI 53705 \\
\texttt{@cs.wisc.edu} \\

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
Financial data is fun.

\section{Initial Data Processing}
Our data set consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 27 parameters ($P$) known to have some affect on the price.
These values are taken at many different time stamps.
Our goal is to predict the stock price in the near future, given current values of the parameters.

First, we create labeled data to use in training our models using the following process:
We sort the data set in increasing time.
For each line, we define $\delta$, the difference in stock price after 2 minutes, to be the current stock price subtracted from the stock price that is associated with the datapoint that is closest to 2 minutes in the future.
(Note that our datapoints are roughly 50 milliseconds apart, so we can expect that the variation in distance between sets of data points will be much less than modeling error.)
We will attempt to build a model which calculates $\delta_t$ when given $P_t$, the values of the 27 parameters at time $t$.

% Note to self: this is done in strategy

\subsubsection*{Acknowledgments}
Thanks to Jerry Zhu for an excellent class, and to Rasmussen and Willams for their excellent book.

\subsubsection*{References}

\small{
}

\end{document}